---
title: "Machine Learning on Barbell Training"
author: "Don VanDemark"
date: "Saturday, June 21, 2014"
output: html_document
---

##Data Loading
```{r}
library(caret)
set.seed(524)
train <- read.csv("./pml-training.csv")
summary(train)
```
As you can see from the summary, there are a large number of variables with mostly blank or NA data. Those won't do us any good so we'll remove them.

```{r}
#remove columns with mostly NA or blank values
train <- subset(train, ,  -c(X,kurtosis_roll_belt,kurtosis_picth_belt,kurtosis_yaw_belt,skewness_roll_belt,skewness_roll_belt.1,skewness_yaw_belt,max_roll_belt,max_picth_belt,max_yaw_belt,min_roll_belt,min_pitch_belt,min_yaw_belt,amplitude_roll_belt,amplitude_pitch_belt,amplitude_yaw_belt,var_total_accel_belt,avg_roll_belt,stddev_roll_belt,var_roll_belt,avg_pitch_belt,stddev_pitch_belt,var_pitch_belt,avg_yaw_belt,stddev_yaw_belt,var_yaw_belt,var_accel_arm,avg_roll_arm,stddev_roll_arm,var_roll_arm,avg_pitch_arm,stddev_pitch_arm,var_pitch_arm,avg_yaw_arm,stddev_yaw_arm,var_yaw_arm,kurtosis_roll_arm,kurtosis_picth_arm,kurtosis_yaw_arm,skewness_roll_arm,skewness_pitch_arm,skewness_yaw_arm,max_roll_arm,max_picth_arm,max_yaw_arm,min_roll_arm,min_pitch_arm,min_yaw_arm,amplitude_roll_arm,amplitude_pitch_arm,amplitude_yaw_arm,kurtosis_roll_dumbbell,kurtosis_yaw_dumbbell,kurtosis_picth_dumbbell,skewness_roll_dumbbell,skewness_yaw_dumbbell,skewness_pitch_dumbbell,max_roll_dumbbell,max_picth_dumbbell,max_yaw_dumbbell,min_roll_dumbbell,min_pitch_dumbbell,min_yaw_dumbbell,min_roll_dumbbell,amplitude_roll_dumbbell,amplitude_pitch_dumbbell,amplitude_yaw_dumbbell,var_accel_dumbbell,avg_roll_dumbbell,stddev_roll_dumbbell,var_roll_dumbbell,avg_pitch_dumbbell,stddev_pitch_dumbbell,var_pitch_dumbbell,avg_yaw_dumbbell,stddev_yaw_dumbbell,var_yaw_dumbbell,kurtosis_roll_forearm,kurtosis_yaw_forearm,kurtosis_picth_forearm,skewness_roll_forearm,skewness_yaw_forearm,skewness_pitch_forearm,max_roll_forearm,max_picth_forearm,max_yaw_forearm,min_roll_forearm,min_pitch_forearm,min_yaw_forearm,amplitude_roll_forearm,amplitude_pitch_forearm,amplitude_yaw_forearm,var_accel_forearm,avg_roll_forearm,stddev_roll_forearm,var_roll_forearm,avg_pitch_forearm,stddev_pitch_forearm,var_pitch_forearm,avg_yaw_forearm,stddev_yaw_forearm,var_yaw_forearm))
```

Now we need to split our training data so we have something to test how accurate our predictor is.
```{r}
inTrain <- createDataPartition(y=train$classe,p=0.75,list=FALSE)
trainSub <- train[inTrain,]
testSub <- train[-inTrain,]
```
We're going to use random forests with a cross-validation. The cross-validation will help reduce the amount of overfitting.
```{r}
forestModel <- train(classe~.,trainSub,importance=TRUE,trControl = trainControl(method="cv",number=2))
```
Before we dive into understanding the model, let's see how well it did.
```{r}
pred<-predict(forestModel,testSub)
confusionMatrix(pred,testSub$classe)
```
The in-sample error rate is 0.0008, which is really good, but we'll expect the out-of-sample error rate to be higher, hopefully only slightly. Interesting that the sensitivity of Class C seems to be the lowest, meaning that it had the hardest time matching that action.

Now let's try to pick apart the model a little.
```{r}
forestModel$results
treeSummary <- getTree(forestModel$finalModel,k=41)
hist(treeSummary[,3],breaks=60,plot=FALSE)
```
First, we have to ignore the fact that zero got an inordinate amount of counts. That just specified the end of a branch. After that, we see that variable 5 and 27 had the most influence on the model. These are "new_window" and "accel_arm_x". Further analysis could be done to see if concentrating on these two gives us a more understandable model.

Finally, let's run the model against the test dataset. This dataset actually had a 100% accuracy rate when run against the grader, however, I would expect a larger test set to actually miss a few and come close to the accuracy of the training set.
```{r}
test <- read.csv("./pml-testing.csv")
test <- subset(test, ,  -c(X,kurtosis_roll_belt,kurtosis_picth_belt,kurtosis_yaw_belt,skewness_roll_belt,skewness_roll_belt.1,skewness_yaw_belt,max_roll_belt,max_picth_belt,max_yaw_belt,min_roll_belt,min_pitch_belt,min_yaw_belt,amplitude_roll_belt,amplitude_pitch_belt,amplitude_yaw_belt,var_total_accel_belt,avg_roll_belt,stddev_roll_belt,var_roll_belt,avg_pitch_belt,stddev_pitch_belt,var_pitch_belt,avg_yaw_belt,stddev_yaw_belt,var_yaw_belt,var_accel_arm,avg_roll_arm,stddev_roll_arm,var_roll_arm,avg_pitch_arm,stddev_pitch_arm,var_pitch_arm,avg_yaw_arm,stddev_yaw_arm,var_yaw_arm,kurtosis_roll_arm,kurtosis_picth_arm,kurtosis_yaw_arm,skewness_roll_arm,skewness_pitch_arm,skewness_yaw_arm,max_roll_arm,max_picth_arm,max_yaw_arm,min_roll_arm,min_pitch_arm,min_yaw_arm,amplitude_roll_arm,amplitude_pitch_arm,amplitude_yaw_arm,kurtosis_roll_dumbbell,kurtosis_yaw_dumbbell,kurtosis_picth_dumbbell,skewness_roll_dumbbell,skewness_yaw_dumbbell,skewness_pitch_dumbbell,max_roll_dumbbell,max_picth_dumbbell,max_yaw_dumbbell,min_roll_dumbbell,min_pitch_dumbbell,min_yaw_dumbbell,min_roll_dumbbell,amplitude_roll_dumbbell,amplitude_pitch_dumbbell,amplitude_yaw_dumbbell,var_accel_dumbbell,avg_roll_dumbbell,stddev_roll_dumbbell,var_roll_dumbbell,avg_pitch_dumbbell,stddev_pitch_dumbbell,var_pitch_dumbbell,avg_yaw_dumbbell,stddev_yaw_dumbbell,var_yaw_dumbbell,kurtosis_roll_forearm,kurtosis_yaw_forearm,kurtosis_picth_forearm,skewness_roll_forearm,skewness_yaw_forearm,skewness_pitch_forearm,max_roll_forearm,max_picth_forearm,max_yaw_forearm,min_roll_forearm,min_pitch_forearm,min_yaw_forearm,amplitude_roll_forearm,amplitude_pitch_forearm,amplitude_yaw_forearm,var_accel_forearm,avg_roll_forearm,stddev_roll_forearm,var_roll_forearm,avg_pitch_forearm,stddev_pitch_forearm,var_pitch_forearm,avg_yaw_forearm,stddev_yaw_forearm,var_yaw_forearm))
predReal <- predict(forestModel,test)
predReal
```
